---
title: "Problem Set 2 (due Week 8, Friday, noon)"
author: 'Daphne Wang (Aditya Maheshwari)'
date: "01/11/2019"
output:
  html_document: default
  pdf_document: default
---

For this problem set, answer the questions in both sections below. Note that each section has an optional question that will not be graded or counted toward this problem set. 

# Section 1: Election Prediction Based on Betting Markets

In the book, class, and tutorials, we studied the prediction of election outcomes using polls. In this section, we study the prediction of election outcomes based on betting markets. In particular, we analyze data for the 2008 US presidential election from an online betting company, called Intrade. At Intrade, people trade contracts such as 'Obama to win the electoral votes of Florida.' Each contract's market price fluctuates based on its sales. Why might we expect betting markets like Intrade to accurately predict the outcomes of elections or of other events? Some argue that the market can aggregate available information efficiently. In this exercise, we will test this *efficient market hypothesis* by analyzing the market prices of contracts for Democratic and Republican nominees' victories in each state. The data file is available in CSV format as `intrade08.csv`. The variables in these datasets are:

 Name          Description
 ------------- --------------------------------------------------------------------
 `day`         Date of the session 
 `statename`   Full name of each state (including District of Columbia in 2008) 
 `state`       Abbreviation of each state (including District of Columbia in 2008) 
 `PriceD`      Closing price (predicted vote share) of Democratic Nominee's market 
 `PriceR`      Closing price (predicted vote share) of Republican Nominee's market 
 `VolumeD`     Total session trades of Democratic Party Nominee's market 
 `VolumeR`     Total session trades of Republican Party Nominee's market 

Each row represents daily trading information about the contracts for either the Democratic or Republican Party nominee's victory in a particular state. We will also use the election outcome data in the file `pres08.csv` with variables:

Name            Description
--------------- ------------------------------------------------
 `state.name`   Full name of state 
 `state`        Two letter state abbreviation 
 `Obama`        Vote percentage for Obama 
 `McCain`       Vote percentage for McCain 
 `EV`           Number of electoral college votes for this state

You may also use poll data from 2008 in the file `polls08.csv. The variables in the polling data are:

Name         Description
------------ -----------------------------------------------------
`state`      Abbreviated name of state in which poll was conducted 
`Obama`      Predicted support for Obama (percentage)
`Romney`     Predicted support for Romney (percentage) 
`Pollster`   Name of organization conducting poll 
`middate`    Middle of the period when poll was conducted 

## Question 1

We will begin by using the market prices on the day before the election to predict the 2008 election outcome. To do this, subset the data such that it contains the market information for each state and candidate only on the day before the election. Note that in 2008 the election day was November 4. 
To get you started, here is the code for loading and merging the data:

```{r}
intrade08<-read.csv("~/Dropbox/cars1/INAF5016/Problem Set 2 due Week 8-20191015/intrade08.csv")
pres08 <- read.csv("~/Dropbox/cars1/INAF5016/Problem Set 2 due Week 8-20191015/pres08.csv")
intresults08 <- merge(intrade08, pres08, by = "state")
# a quick look at the merged data
head(intresults08)
marketinfo<-intresults08[intresults08$day=="2008-11-03",]
```
We compare the closing prices for the two candidates in a given state and classify a candidate whose contract has a higher price as the predicted winner of that state. Which states were misclassified? (It is sufficient to only report the misclassified states.) How does this compare to the classification by polls presented in the chapter? 


```{r}
miscalculated<-0
for (i in marketinfo$statename) {
  m<- subset(marketinfo, statename == i)
  if ((m$PriceD>m$PriceR&m$Obama<m$McCain)|(m$PriceD<m$PriceR&m$Obama>m$McCain)){
miscalculated[i]<-m$statename
  } 
}
miscalculated
```
In the polls classifaction presented in the chapter, there are three misclassified states as opposed to two found in this analysis.

## Question 2

How do the predictions based on the betting markets change over time? Use the classification procedure as above on each of the last 90 days of the 2008 campaign rather than just the day before the election. (Hint: Use a loop.) Plot the predicted number of electoral votes for the Democratic party nominee over this 90-day period. The resulting plot should also indicate the day of the election and the actual election result. (Hint: Use `abline()` for these.) Note that in 2008, Obama won 365 electoral votes. Briefly comment on the plot.

```{r}
#create last days vector that counts number of days to election
intresults08$lastdays<-as.Date("2008-11-04")-as.Date(intresults08$day)

#create sequence of last ninety days
x<-as.Date("2008-11-03")
x-89
y<-as.Date("2008-08-06")
xy<-seq(y,x, by = 1)
z<-seq(1:90)
# define empty container for 90 days to run loop through
dem.votes<-rep(NA,90)
names(dem.votes)<-rev(xy)

#calculate number of electoral votes per day for each of the 90 days leading up to election
for (i in z){
  n<-subset(intresults08, lastdays==i)
  PD<- na.omit(sum(ifelse(n$PriceD>n$PriceR,n$EV,0)))
  dem.votes[i]<-PD}
dem.votes
```

```{r}
#calculate actual election result based on Obama and McCain's margin 
i <- 0
n<-subset(intresults08, lastdays==i)
realmargin<- sum(ifelse(n$Obama>n$McCain,n$EV,0))
realmargin

plot(xy,rev(dem.votes), xlim=c(as.Date("2008-08-04"),as.Date("2008-11-04")), xlab="date", ylab="predicted votes", main="Number of Predicted Electoral Votes 90 Days Before Election")
    text(as.Date("2008-09-24"),370, "Actual Election Result")
    text(as.Date("2008-10-24"),270, "Election Date")
     abline(v=as.Date("2008-11-04"))
     abline(h=365) 
```

The plot shows that the predicted number of votes that Obama wins get closer to the actual number of votes as the days get closer to Election Day. It is also important to note that Obama actually won 365 electoral votes but according to the margin between Obama and McCain's margin, Obama would win 364 votes.     


## Question 3

Repeat the previous exercise but this time use the seven-day *moving-average* price, instead of the daily price, for each candidate within a state. (Hint: This can be done by recycling code from the previous answer, with some changes inside the loop.) For a given day, we take the average of the Session Close prices within the past seven days (including that day). To answer this question, we must first compute the seven-day average within each state. Next, we sum the electoral votes for the states Obama is predicted to win. Using the `tapply()` function will allow us to efficiently compute the predicted winner for each state on a given day.

```{r}
# define empty container for 90 days to run loop through for moving average price
mdem.votes<-rep(NA,90)
names(mdem.votes)<-rev(xy) 

 for (i in 1:90) {
selected.day.7<-subset(intresults08, subset=((lastdays<=(90-i+7)) & (lastdays>=(90-i+1))))
meanD<-tapply(selected.day.7$PriceD, selected.day.7$state, mean)
meanR<-tapply(selected.day.7$PriceR, selected.day.7$state, mean)

uEV<-tapply(selected.day.7$EV,selected.day.7$state,unique)
margin<-data.frame(meanD, meanR, uEV)
row.names(margin)<-1:51
Obamawin<- sum(ifelse(margin$meanD>margin$meanR,margin$uEV,0))
mdem.votes[i]<-Obamawin
 }

# number of electoral votes based on moving-average price
mdem.votes

plot(xy,mdem.votes, xlim=c(as.Date("2008-08-04"),as.Date("2008-11-04")), type="p", xlab="date", col="purple", ylab="predicted votes", main="Number of Predicted Electoral Votes 90 Days Before Election")
    text(as.Date("2008-09-24"),362, "Actual Election Result")
    text(as.Date("2008-10-26"),280, "Election Date")
     abline(v=as.Date("2008-11-04"))
     abline(h=365) 
```

The moving average price is able to give a more accurate prediction earlier.   

## Question 4 (OPTIONAL, if you just can't get enough of loops)

Create a similar plot for 2008 state-wide poll predictions using the data file `polls08.csv`. Notice that polls are not conducted daily within each state. Therefore, within a given state for each of the last 90 days of the campaign, we compute the average margin of victory from the most recent poll(s) conducted. If multiple polls occurred on the same day, average these polls. Based on the most recent predictions in each state, sum Obama's total number of predicted electoral votes. One strategy to answer this question is to program two loops - an inner loop with 51 iterations for each state and an outer loop with 90 iterations for each day. (Hint: You can use a counter in the inner loop to collect electoral votes per state, and in the outer loop save/assign the result and reset the counter.)

## Question 5

What is the relationship between the price margins of the Intrade market and the actual margin of victory? 

Using only the market data from the day before the election in 2008, regress Obama's actual margin of victory in each state on Obama's price margin from the Intrade markets. 

Considering only the trading one day from the election, predict the actual electoral margins from the trading margins using a linear model. Does it predict well? How would you visualize the predictions and the outcomes together? (Hint: Because we only have one predictor you can use `abline()`.) Similarly, in a separate analysis, regress Obama's actual margin of victory on the Obama's predicted margin from the latest polls within each state. Interpret the results of these regressions.

```{r}
# margin in trade prices
marketinfo$trade.diff<- marketinfo$PriceD-marketinfo$PriceR
# actual electoral margins
marketinfo$actual.diff<- marketinfo$Obama-marketinfo$McCain

#predict the actual electoral margins from the trading margins using a linear model
st88<-rep(NA,51)
names(st88)<-marketinfo$state
for (i in 1:51){
fitbit<-subset(marketinfo, state=i)
fit <- lm(actual.diff ~ trade.diff, data = fitbit)
st88[i]<-coef(fit)["trade.diff"] }
fit

#predict electoral votes based on price margins
sum(ifelse(marketinfo$trade.diff>0, uEV, 0))

#visualize the predictions and the outcomes together
plot(marketinfo$trade.diff, marketinfo$actual.diff, type = "p", xlab="predicted margin", ylab="actual margin", main="Predicted vs Actual Margin of Victory")
abline(fit)
```

The linear regression model predicts the relationship between predicted and actual margins well as most points fall close to the line. The trade price margins predict that Obama would win 364 votes and he actually won 365. When the margins are predicted to be negative, the actual margins are also negative. The converse is true for positive values. For the outlier states in the top right corner, it shows that the prediction is right in that obama won. However, the margin by which he won by was predicted to be lower. 
```{r}
# calculate predicted margin of victory
polls08 <- read.csv("~/Dropbox/cars1/INAF5016/Problem Set 2 due Week 8-20191015/polls08.csv")
polls08$margin<- polls08$Obama-polls08$McCain
polls08$middate <- as.Date(polls08$middate)
polls08$DaysToElection <- as.Date("2008-11-04") - polls08$middate

#calculate actual margin of victory
pres08$margin<-pres08$Obama-pres08$McCain

#create empty container for predicted margin of victory from latest polls within each state
poll.pred <- rep(NA, 51) 
st.names <- unique(polls08$state)
names(poll.pred) <- as.character(st.names)

#find margin of victory from latest polls within each state
for (i in 1:51){
state.data <- subset(polls08, subset = (state == st.names[i]))
latest <- subset(state.data, DaysToElection == min(DaysToElection))
poll.pred[i]<- mean(latest$margin)
}

#regress actual margin on predicted margin from latest polls within each state
actual.vs.poll<- lm(pres08$margin~poll.pred)
actual.vs.poll
```

Both regressions show that the predicted margin of victory is a pretty accurate representation of the actual margin of victory. In both cases, the intercept is close to zero and the direction of the slope is positive. This shows that if the predicted margins are negative or to the left of the intercept, (meaning Obama is predicted to lose the state), the actual results would also be likely to be negative.

## Question 6 

Even efficient markets are not omniscient. Information comes in about the election every day and the market prices should reflect any change in information that seem to matter to the outcome. We can examine how and about what the markets change their minds by looking at which states they are confident about, and which they update their ‘opinions’ (i.e. their prices) about. Over the period before the election, let’s see how prices for each state are evolving. We can get a compact summary of price movement by fitting a linear model to Obama’s margin for each state over the 20 days before the election. We will summarise price movement by the direction (up or down) and rate of change (large or small) of price over time. (This is basically also what people in finance do.)

Start by plotting Obama’s margin in West Virginia against the number of days until the election and modeling the relationship with a linear model. (Hint: West Virginia is 50th on the alphabetical list of state names.) Use the last 20 days. 
Show the model's predictions on each day and the data. What does this model's slope coefficient tells us about which direction the margin is changing and also how fast it is changing? 
Then do it for all states and collect the slope coefficients to see how volatile the state estimates are. Show the distribution of these slopes with a histogram.
```{r}
# subset for West Virginia predictions 20 days before election
last.20<-subset(intresults08, subset=((lastdays<=20) & (lastdays>=1)))
last.20$lastdays<--last.20$lastdays
last.20$pricemargin<-last.20$PriceD-last.20$PriceR
WVlast.20<-last.20[last.20$state=="WV",]


#regress price margin against last 20 days in West Viginia
WVlm<-lm(WVlast.20$pricemargin~WVlast.20$lastdays)
WVlm
#predictions on each day
predictx<-data.frame(lastdays=WVlast.20$lastdays)
werk<-predict(WVlm, newdata=predictx)
werk

#plot West Virginia's price margin in the 20 days before election
plot(WVlast.20$lastdays, WVlast.20$pricemargin, xlab="days til election", ylab="price margin", main="West Virginia Price Margin 20 days before Election")
abline(WVlm)
```

The slope coefficient tells us that the margin is changing in a negative direction at -3.41 price per day. In other words, the confidence that Obama would win in West Virginia is decreasing by 3.41 points every day as we approach the election date.

```{r}
# create empty container for 51 states
results<-rep(NA,51)
names(results)<-unique(last.20$state)

#find slope coefficients for every states' regression of price margin against the 20 days leading to election 
for (i in last.20$state){
  st8<-subset(last.20, state == i)
  st.coeff<- (lm(pricemargin~lastdays, data=st8))
  st.slope<-coef(st.coeff)[2] 
  results[i]<-st.slope}
results
hist(results, freq = FALSE, breaks = seq(from = -4, to = 4, by = 0.2))
```

Unlike West Virginia, where the slope coefficient is -3.44, most states' slopes are close to 0, this shows that the prices are not very volatile for the last 20 days. Slopes that are closer to 0 shows that the price margins are not changing much within the 20 day period. Also, there seems to be a fairl equal number of positive and negative slope coefficients which shows that there are some states that are growing more confident towards voting for Obama and others that are growing less confident towards voting for Obama.

## Question 7

Now predict the winner of the election *one week before the election* using the Intrade data. To do so, first use the *two weeks before* that moment to fit state level linear models, then use those models to predict what will happen in each state. How well does the model do predicting the election outcome?


```{r}
# subset for 
twowb<-subset(intresults08, subset=((lastdays<=21) & (lastdays>=8)))
twowb$pricemargin<-twowb$PriceD-twowb$PriceR
st.EV<-tapply(twowb$EV, twowb$state, unique)

wb.results<-rep(NA,51)
names(wb.results)<-unique(last.20$state)


no.days.to.election <- as.Date("2008-11-04") - as.Date("2008-10-28")

for (i in twowb$state){
  twowb.st8<-subset(twowb, state == i)
  twowb.lm<- (lm(pricemargin~lastdays, data=twowb.st8))
  predict.2wb<-predict(twowb.lm, newdata=data.frame(lastdays = no.days.to.election))
  wb.results[i]<-predict.2wb
  }
wb.results
#predicted number of electoral votes for Obama
sum(ifelse(wb.results>0, st.EV, 0))

```
This model overpredicts the number of electoral votes that Obama would win. Using the predicted vote margins based on the price margins in week 2 and 3, it looks like Obama would win 375 electoral votes on election day. This is more than the 365 votes that he actually received. 

# Section 2: Immigration Attitudes

Why do the majority of voters in the U.S. and other developed countries oppose increased immigration? According to the conventional wisdom and many economic theories, people simply do not want to face additional competition on the labor market (*economic threat* hypothesis). Nonetheless, most comprehensive empirical tests have failed to confirm this hypothesis and it appears that people often support policies that are against their personal economic interest. At the same time, there has been growing evidence that immigration attitudes are rather influenced by various deep-rooted ethnic and cultural stereotypes (*cultural threat* hypothesis). Given the prominence of workers' economic concerns in the political discourse, how can these findings be reconciled?

This exercise is based in part on Malhotra, N., Margalit, Y. and Mo, C.H., 2013. "[Economic Explanations for Opposition to Immigration: Distinguishing between Prevalence and Conditional Impact](https://dx.doi.org/10.1111/ajps.12012)." *American Journal of Political Science*, Vol. 38, No. 3, pp. 393-433.

The authors argue that, while job competition is not a prevalent threat and therefore may not be detected by aggregating survey responses, its *conditional* impact in selected industries may be quite sizable. To test their hypothesis, they conduct a unique survey of Americans' attitudes toward H-1B visas. The plurality of H-1B visas are occupied by Indian immigrants, who are skilled but ethnically distinct, which enables the authors to measure a specific skill set (high technology) that is threatened by a particular type of immigrant (H-1B visa holders). The data set `immig.csv` has the following variables:

 Name                Description
 ------------------- --------------------------------------------------------------------
 `age`               Age (in years)
 `female`            `1` indicates female; `0` indicates male
 `employed`          `1` indicates employed; `0` indicates unemployed
 `nontech.whitcol`   `1` indicates non-tech white-collar work (e.g., law)
 `tech.whitcol`      `1` indicates high-technology work
 `expl.prejud`       Explicit negative stereotypes about Indians (continuous scale, 0-1)
 `impl.prejud`       Implicit bias against Indian Americans (continuous scale, 0-1)
 `h1bvis.supp`       Support for increasing H-1B visas (5-point scale, 0-1)
 `indimm.supp`       Support for increasing Indian immigration (5-point scale, 0-1)

The main outcome of interest (`h1bvis.supp`) was measured as a following survey item: "Some people have proposed that the U.S. government should increase the number of H-1B visas, which are allowances for U.S. companies to hire workers from foreign countries to work in highly skilled occupations (such as engineering, computer programming, and high-technology). Do you think the U.S. should increase, decrease, or keep about the same number of H-1B visas?" Another outcome (`indimm.supp`) similarly asked about the "the number of immigrants from India." Both variables have the following response options: `0` = "decrease a great deal", `0.25` = "decrease a little", `0.5` = "keep about the same", `0.75` = "increase a little", `1` = "increase a great deal".

To measure explicit stereotypes (`expl.prejud`), respondents were asked to evaluate Indians on a series of traits: capable, polite, hardworking, hygienic, and trustworthy. All responses were then used to create a scale lying between `0` (only positive traits of Indians) to `1` (no positive traits of Indians). Implicit bias (`impl.prejud`) is measured via the *Implicit Association Test* (IAT) which is an experimental method designed to gauge the strength of associations linking social categories (e.g., European vs Indian American) to evaluative anchors (e.g., good vs bad). Individual who are prejudiced against Indians should be quicker at making classifications of faces and words when *European American* (*Indian American*) is paired with *good* (*bad*) than when *European American* (*Indian American*) is paired with *bad* (*good*). 
<!-- If you want, you can test yourself [here](https://implicit.harvard.edu/implicit/takeatest.html). -->


## Question 1

Start by examining the distribution of immigration attitudes (as factor variables). What is the proportion of people who are willing to increase the quota for high-skilled foreign professionals (`h1bvis.supp`) or support immigration from India (`indimm.supp`)?

```{r}
immig<-read.csv("~/Dropbox/cars1/INAF5016/Problem Set 2 due Week 8-20191015/immig.csv")

immig$h1bvis.suppx<-as.factor(immig$h1bvis.supp)
lev <- levels(immig$h1bvis.suppx) 
lev

#define those who are willing to increase the quota for high-skilled foreign professionals, as those who scored 0.75 or 1 on h1bvis.supp 
immig$h1bvis.inc <- 
    ifelse(immig$h1bvis.supp == lev[4] | immig$h1bvis.supp == lev[5], 1, 0)
## proportion of people who are willing to increase the quota for high-skilled foreign professionals
sum(immig$h1bvis.inc, na.rm=TRUE)/ length(immig$h1bvis.inc)

#define those who support immigration, as those who scored 0.5,  0.75, or 1 on h1bvis.supp 
immig$indimm.inc <- 
    ifelse(immig$indimm.supp == lev[3]|immig$indimm.supp == lev[4] | immig$indimm.supp == lev[5], 1, 0)

#proportion of people who are support immigration
sum(immig$indimm.inc, na.rm=TRUE)/ length(immig$indimm.inc)
immig$h1bvis.supp<-as.numeric(immig$h1bvis.supp)

```
Now compare the distribution of two distinct measures of cultural threat: explicit stereotyping about Indians (`expl.prejud`) and implicit bias against Indian Americans (`impl.prejud`). In particular, create a scatterplot, add a linear regression line to it, and calculate the correlation coefficient. Based on these results, what can you say about their relationship?
```{r}
cult.sub<-na.omit(data.frame(immig$expl.prejud,immig$impl.prejud))
colnames(cult.sub)<-c("expl", "impl")

plot(cult.sub$impl,cult.sub$expl, xlab="implicit prejudice", ylab="explicit prejudice", main="Implicit vs Explicit Prejudice")
cult<-lm(cult.sub$expl~cult.sub$impl)
abline(cult)
cult
cor(cult.sub$impl,cult.sub$expl)
```

There is no clear relationship between explicit stereotyping and implicit bias. The slope coefficient is close to 0 which shows that an increase in explicit stereotyping does not necessarily show an increase in implicit bias. Also, the correlation is close to zero which does not indicate a strong correlation. 

## Question 2

Compute the correlations between all four policy attitude and cultural threat measures. Do you agree that cultural threat is an important predictor of immigration attitudes as claimed in the literature?
If the labor market hypothesis is correct, opposition to H-1B visas should also be more pronounced among those who are economically threatened by this policy such as individuals in the high-technology sector. At the same time, tech workers should not be more or less opposed to general Indian immigration because of any *economic* considerations. First, regress H-1B and Indian immigration attitudes separately on the indicator variable for tech workers (`tech.whitcol`). Do the results support the hypothesis? Is the relationship different from the one involving cultural threat and, if so, how?

```{r}
immigvariables<-data.frame(immig$expl.prejud,immig$impl.prejud, immig$indimm.supp, immig$h1bvis.supp)
cor(immigvariables, use="complete.obs")

```
The results show that immigration and h1b visa support show a high positive correlation of 0.6107. This means that support for one would often show support for another. However, there is a weak negative correlation between all other policy attitudes and cultural threat measures except for the correlation between immigration support and explicit prejudice which shows a negative correlation of -0.321. This means that the immigration support shows a moderately inverse relationship with explicit prejudice. 


```{r}
cult.sub<-na.omit(data.frame(immig$indimm.supp, immig$h1bvis.supp, immig$tech.whitcol))
colnames(cult.sub)<-c("indimm", "h1bvis", "whitcol")
immig$indimm<- as.numeric(immig$indimm.supp)
immig$h1bvis<- as.numeric(immig$h1bvis.supp)
immig$tech.whitcol<-as.numeric(immig$tech.whitcol)
lm(immig$h1bvis~immig$tech.whitcol)
cor(immig$tech.whitcol,immig$h1bvis, use="complete.obs")

lm(immig$indimm~immig$tech.whitcol)
cor(immig$tech.whitcol,immig$indimm, use="complete.obs")
```
The results do not support the hypothesis because the correlation is very low and weak in both comparisons. As well, the slope coefficient for both regressions are very low which means that there is no clear relationship between tech workers and immigration support or tech immigration and h1b visa support. This disproves the hypothesis because it shows that tech workers do not show a clear opposion for the h1b visa. Cultural threat shows a stronger correlation with immigation opposition than economic threat.

## Question 3

When examining hypotheses, it is always important to have an appropriate comparison group. One may argue that comparing tech workers to everybody else as we did in Question 2 may be problematic due to a variety of confounding variables (such as skill level and employment status). First, create a single factor variable `group` which takes a value of `tech` if someone is employed in tech, `whitecollar` if someone is employed in other "white-collar" jobs (such as law or finance), `other` if someone is employed in any other sector, and `unemployed` if someone is unemployed. Then, compare the support for H-1B across these conditions by using linear regression. Interpret the results: is this comparison more or less supportive of the labor market hypothesis than the one in Question 2?
```{r}
immig$group<-c(ifelse(immig$employed==0,"unemployed",(ifelse(immig$tech.whitcol==1,"tech", (ifelse(immig$nontech.whitcol==1,"whitecollar","other"))))))
immig$group<-as.factor(immig$group)

lm(immig$h1bvis.supp~-1+immig$group)

```
The correlation between for H-1B support and working grouup is positive for all groups. This shows that all groups support the H-1B to some degree. The group with the weakest correlation for H-1B visa support or the group with the lowest support is the tech worker group. This comparison provides better support for the labor market hypothesis than the one in Question 2 because it shows that tech workers show the least support for the H1-B visa.

Now, one may also argue that those who work in the tech sector are disproportionately young and male which may confound our results. To account for this possibility, fit another linear regression but also include `age` and `female` as pre-treatment covariates (in addition to `group`). Does it change the results and, if so, how?
```{r}
lm(h1bvis.supp~-1+group+age+female, data=immig)
```
After accounting for age and gender as confounding variables, the support for the H1B visa in all groups increases. Gender seems to have a larger effect because it's coefficient is -0.075 and age's coefficient is -0.00248. 

Finally, fit a linear regression model with all threat indicators (`group`, `expl.prejud`, `impl.prejud`) and calculate its $R^2$. How much of the variation is explained? Based on the model fit, what can you conclude about the role of threat factors?
```{r}
allfit<-lm(h1bvis.supp~-1+group+expl.prejud+impl.prejud, data=immig)
allfit


wrsg<-summary(allfit)
wrsg$r.squared
```
Roughly half of the variation is explained From the calculated R^2. Cultural threat factors (explicit and implicit prejudice) seem to have a negative effect on support for the H1B visa support. When taking cultural threat into account as confounding variables, the support for the H1B visa for all the different groups increased. 

## Question 4

Besides economic and cultural threat, many scholars also argue that gender is an important predictor of immigration attitudes. While there is some evidence that women are slightly less opposed to immigration than men, it may also be true that gender conditions the very effect of other factors such as cultural threat. To see if it is indeed the case, fit a linear regression of H-1B support on the interaction between gender and implicit prejudice. Then, create a plot with the predicted level of H-1B support (y-axis) across the range of implicit bias (x-axis) by gender. Considering the results, would you agree that gender alters the relationship between cultural threat and immigration attitudes?
```{r}
fit<-lm(h1bvis.supp~female*impl.prejud, data=immig)
fit

#data frame for female and male predictions
gurl<-predict(fit, newdata = data.frame(female = 1, impl.prejud = immig$impl.prejud))
boi<-predict(fit, newdata = data.frame(female = 0, impl.prejud = immig$impl.prejud))
gurl
boi
plot(immig$impl.prejud, gurl, type = "l",
     ylim=c(0.2,0.6),
xlab = "Implicit Bias", ylab = "H1b Visa Support", main="Implicit Prejudice vs. H-1B Visa Support")
lines(immig$impl.prejud, boi, col="green")
text(4, 6, "Average effect on PRI support")
text(4, -4, "Average effect on turnout")
legend(x = 0, y = 0.3, bty = "n",
legend = c("female", "male"),
col = c("black", "green"), lty = 1)
```

Gender alters the relationship between cultural threat and immigration attitudes. The model above shows that males who have a higher implicit bias are predicted to have lower support for the H1B visa than females and males who have a lower bias are predicted to have a higher support for the H1B visa than females. This shows that there is a stronger relationship between implicit bias and visa support for males than females.    

Age is another important covariate. Fit two regression models in which H-1B support is either a linear or quadratic function of age. Compare the results of both models by plotting the predicted levels of support (y-axis) across the whole age range (x-axis). Would you say that people become more opposed to immigration with age?
```{r}
#linear model of H-1B support across age
al.fit<-lm(h1bvis.supp~age, data=immig)
al.fit
alfit.pred<-predict(al.fit, newdata = data.frame(age=unique(immig$age)))

#quadratic model of H-1B support across age
aq.fit<-lm(h1bvis.supp~I(age^2)+age, data = immig)
aqfit.pred<-predict(aq.fit, newdata = data.frame(age=19:90))

plot(unique(immig$age), alfit.pred, type = "l",
        ylim=c(0.2,0.6),
xlab = "Age", ylab = "H1b Visa Support", main="H-1B Visa Support across Age")
lines(19:90, aqfit.pred, col="green")
text(4, 6, "Average effect on PRI support")
text(4, -4, "Average effect on turnout")
legend(x = 16, y = 0.3, bty = "n",
legend = c("linear model", "quadratic model"),
col = c("black", "green"), lty = 1)
```

The linear model shows that support for the H1B Visa decreases with age. However, the quadratic model shows that visa support decreases with age up until 50 and then it starts increasing again. This could signify that retirees who are not in job competition with immigrants show more support for the visa.

## Question 5 (OPTIONAL, if you're interested)

To corroborate your conclusions with regard to cultural threat, create separate binary variables for both prejudice indicators based on their median value (`1` if `>` than the median) and then compare average H-1B and Indian immigration attitudes (as numeric variables) depending on whether someone is implicitly or explicitly prejudiced (or both). What do these comparisons say about the role of cultural threat?

What about the role of economic threat? One may argue that tech workers are simply more or less prejudiced against Indians than others. To account for this possibility, investigate whether economic threat is in fact distinguishable from cultural threat as defined in the study. In particular, compare the distribution of cultural threat indicator variable using the Q-Q plot depending on whether someone is in the high-technology sector. Would you conclude that cultural and economic threat are really distinct?

```{r}
#binary variable for implicit and explicit prejudice
immig$impl.prejud.bin<-ifelse(immig$impl.prejud>=0.5,1,0)
immig$expl.prejud.bin<-ifelse(immig$expl.prejud>=0.5,1,0)
immig$impl.prejud.bin<- as.numeric(immig$impl.prejud.bin)
immig$expl.prejud.bin<- as.numeric(immig$expl.prejud.bin)
#average H1B visa support for people who are implicitly biased
tapply(immig$h1bvis.supp, immig$impl.prejud.bin, mean)
#average H1B visa support for people who are explicitly biased
tapply(immig$h1bvis.supp, immig$expl.prejud.bin, mean)
#average H1B visa support for people who are implicitly and explicitlt biased
tapply(immig$h1bvis.supp, immig$expl.prejud.bin==1&immig$impl.prejud.bin==1, mean)

#average immigration support for people who are implicitly biased
tapply(immig$indimm.supp, immig$impl.prejud.bin, mean)
#average H1B visa support for people who are explicitly biased
tapply(immig$indimm.supp, immig$expl.prejud.bin, mean)
#average H1B visa support for people who are implicitly and explicitlt biased
tapply(immig$indimm.supp, immig$expl.prejud.bin==1&immig$impl.prejud.bin==1, mean)


tech<-na.omit(immig[immig$tech.whitcol==1,])
nontech<-na.omit(immig[immig$tech.whitcol==0,])
qqplot(tech$expl.prejud, tech$impl.prejud, xlab = "Explicit Prejudice", ylab = "Implicit Prejudice", main="Prejudice for Tech Workers")
abline(0,1)
qqplot(nontech$expl.prejud, nontech$impl.prejud, xlab = "Explicit Prejudice", ylab = "Implicit Prejudice", main="Prejudice for Non-Tech Workers")
abline(0,1)
```
Both tech and non-tech workers show higher implicit than explicit prejudice. This shows that cultural and economic threat are not very distinct. 